---
title: "Data605 - Assignment5"
author: "Kishore Prasad"
date: "September 25, 2016"
output: html_document
---


##Problem set 1


In this problem set we'll work out some properties of the least squares solution that we reviewed in the weekly readings. Consider the unsolvable system Ax = b as given below:

$\begin{bmatrix} 1 & 0 \\ 1 & 1 \\ 1 & 3 \\ 1 & 4  \end{bmatrix} \begin{bmatrix} x_1 \\ x_2 \end{bmatrix} = \begin{bmatrix} 0 \\ 8 \\ 8 \\ 20  \end{bmatrix}$


**Answer:**

####a. Write R markdown script to compute ATA and ATb.

Lets first define the 2 matrices. 

```{r, warning=FALSE, message=FALSE}
# Define the matrix
A <- matrix(c(1,1,1,1,0,1,3,4), nrow = 4, ncol = 2, byrow = FALSE)
A

b <- matrix(c(0,8,8,20), nrow = 4, ncol = 1, byrow = FALSE)
b

```


We then calculate 

$l = A^T A$ 

```{r, warning=FALSE, message=FALSE}
l <- t(A) %*% A
l
```

and 

$m = A^T b$.

```{r, warning=FALSE, message=FALSE}
m <- t(A) %*% b
m
```


####b. Solve for ^x in R using the above two computed matrices.

We know that:

$A^TA \times \hat { x }  = A^T b$ 

Hence it follows that, 

$\hat { x }  = (A^TA)^{-1} \times A^Tb$

i.e 

$\hat { x }  = l^{-1} \times m$


```{r, warning=FALSE, message=FALSE}
x_hat <- solve(l) %*% m
x_hat
```

####c. What is the squared error of this solution?

We know that $b = p + e$, where p is the projection vector and e is the error vector. To compute the error, we can rewrite as $e = b - p$.

To calculate p, we use the below:

$p = A \times \hat{x}$

```{r, warning=FALSE, message=FALSE}
projection <- A %*% x_hat
projection
```

We now calculate squared error as below:

```{r, warning=FALSE, message=FALSE}
e <- b - projection
e

squared_error_b <- sum(e^2)
squared_error_b
```


####d. Instead of b = [0; 8; 8; 20], start with p = [1; 5; 13; 17] and find the exact solution (i.e. show that this system is solvable as all equations are consistent with each other. This should result in an error vector e = 0).

Repeating the same steps as above with p = [1, 5, 13, 17]


```{r, warning=FALSE, message=FALSE}
p <- matrix(c(1,5,13,17), nrow = 4, ncol = 1, byrow = FALSE)
p

l <- t(A) %*% A
m <- t(A) %*% p

x_hat <- solve(l) %*% m
x_hat

projection_2 <- A %*% x_hat
projection_2

e <- round(p - projection_2)
e

squared_error_p <- round(sum(e^2))
squared_error_p
```


####e. Show that the error e = b - p = [-1, 3, -5, 3]


```{r, warning=FALSE, message=FALSE}
e <- b - p
e
```


####f. Show that the error e is orthogonal to p and to each of the columns of A. 

If 2 vectors are orthogonal, then their dot product is 0.

Checking for p :

```{r, warning=FALSE, message=FALSE}
dot_e_p <- round(sum(e*p))
dot_e_p
```

Checking for first column of A :

```{r, warning=FALSE, message=FALSE}
dot_e_A1 <- round(sum(e*A[,1]))
dot_e_A1
```

Checking for second column of A :

```{r, warning=FALSE, message=FALSE}
dot_e_A2 <- round(sum(e*A[,2]))
dot_e_A2
```




##Problem set 2

Consider the modified auto-mpg data (obtained from the UC Irvine Machine Learning dataset). This dataset contains 5 columns: displacement, horsepower, weight, acceleration, mpg. We are going to model mpg as a function of the other four variables.

**Answer:**


####a. Write an R markdown script that takes in the auto-mpg data, extracts an A matrix from the first 4 columns and b vector from the fifth (mpg) column.


First, read the data as a dataset.

```{r, warning=FALSE, message=FALSE}
data <- read.table('auto-mpg.data', col.names=c('displacement', 'horsepower', 'weight', 'acceleration', 'mpg'))

head(data)
```

Next, create matrices from the data. 

```{r, warning=FALSE, message=FALSE}

A <- data.matrix(data[,1:4])

# An additional column for the intercepts
A <- cbind(rep(1, nrow(A)), A)
head(A)

b <- data.matrix(data[,5])
head(b)

```

####b. Using the least squares approach, your code should compute the best fitting solution. That is, find the best fitting equation that expresses mpg in terms of the other 4 variables. 

Next, compute the best fitting solution using the same approach as the previous problem set. The resulting vector will give the slope and the parameters for displacement, horsepower, weight and acceleration.

```{r, warning=FALSE, message=FALSE}

l <- t(A) %*% A
m <- t(A) %*% b

x_hat <- solve(l) %*% m
x_hat

```

The resulting equation is given as below : 

$\hat{mpg} = `r x_hat[1]` + (`r x_hat[2]`) * displacement + (`r x_hat[3]`) * horsepower + (`r x_hat[4]`) * weight + (`r x_hat[5]`) * acceleration$ 

####c. Finally, calculate the fitting error between the predicted mpg of your model and the actual mpg. 
```{r, warning=FALSE, message=FALSE}

p <- A %*% x_hat
e <- b - p
squared_error <- sum(e^2)

squared_error
```
