---
title: "Data605 - Assignment1"
author: "Kishore Prasad"
date: "August 29, 2016"
output: html_document
---


1. Problem set 1

You can think of vectors representing many dimensions of related information. For instance, Netflix might store all the ratings a user gives to movies in a vector. This is clearly a vector of very large dimensions (in the millions) and very sparse as the user might have rated only a few movies. Similarly, Amazon might store the items purchased by a user in a vector, with each slot or dimension representing a unique product and the value of the slot, the number of such items the user bought. One task that is frequently done in these settings is to find similarities between users. And, we can use dot-product between vectors to do just that. As you know, the dot-product is proportional to the length of two vectors and to the angle between them. In fact, the dot-product between two vectors, normalized by their lengths is called as the cosine distance and is frequently used in recommendation engines.

```{r, warning=FALSE, message=FALSE, results='asis'}
u <- c(0.5, 0.5)
v <- c(3, -4)
```


(1) Calculate the dot product u.v where u = [0.5; 0.5] and v = [3; -4]

**Answer:**  

```{r, warning=FALSE, message=FALSE, results='asis'}

# short answer
sum(u*v)


#long answer
u[1]*v[1] + u[2]*v[2]

```


(2) What are the lengths of u and v? Please note that the mathematical notion of the length of a vector is not the same as a computer science defnition.

**Answer:** 

```{r, warning=FALSE, message=FALSE, results='asis'}

# short answer
sqrt(sum(u*u))
sqrt(sum(v*v))


#long answer
sqrt(u[1]*u[1] + u[2]*u[2])
sqrt(v[1]*v[1] + v[2]*v[2])



```






(3) What is the linear combination: 3u - 2v?

**Answer:**

```{r, warning=FALSE, message=FALSE, results='asis'}

# short answer
(3*u - 2*v)

#long answer
c((3*u[1] - 2*v[1]), (3*u[2] - 2*v[2])) 

```





(4) What is the angle between u and v

**Answer:**

```{r, warning=FALSE, message=FALSE, results='asis'}

# short answer
# angle in radians
acos(sum(u*v) / (sqrt(sum(u*u)) * sqrt(sum(v*v))))

# angle in degrees
(acos(sum(u*v) / (sqrt(sum(u*u)) * sqrt(sum(v*v)))) * 180) / pi


#long answer
# angle in radians
acos((u[1]*v[1] + u[2]*v[2]) / ((sqrt(u[1]*u[1] + u[2]*u[2])) * (sqrt(v[1]*v[1] + v[2]*v[2])) ) )

# angle in degrees
(acos((u[1]*v[1] + u[2]*v[2]) / ((sqrt(u[1]*u[1] + u[2]*u[2])) * (sqrt(v[1]*v[1] + v[2]*v[2])) ) ) * 180) / pi

```


2. Problem set 2
Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. You don't have to worry about degenerate cases and can safely assume that the function will only be tested with a system of equations that has a solution. Please note that you do have to worry about zero pivots, though. Please note that you should not use the built-in function solve to solve this system or use matrix inverses. The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure. Please test it with the system below and it should produce a solution x = [-1.55, -0.32, 0.95]

```{r, warning=FALSE, message=FALSE, results='asis'}

solve_eq <- function(a, b) {

    # create a vector for storing the x2, x2 and x3 values. 
    x<-c(0,0,0)

    # Augment the matrix and vector to ease operations.
    aug_mat <- cbind(a,b)

        
    # Eliminate x1 in R2 and R3
    aug_mat[2,] <- aug_mat[2,] - ((aug_mat[2,1]/aug_mat[1,1])*aug_mat[1,])
    aug_mat[3,] <- aug_mat[3,] - ((aug_mat[3,1]/aug_mat[1,1])*aug_mat[1,])

    # check for zero pivot in R2
    if(aug_mat[2,2]==0) {
        # check for zero pivot in R3
        if(aug_mat[3,2]==0){
            # if R2 and R3 both contain zero pivots then row exchange does not work. So return without any solution
            return(x)
        }
        else {
            # exchange R2 and R3
            tmp <- aug_mat[2,]
            aug_mat[2,] <- aug_mat[3,]
            aug_mat[3,] <- tmp
        }
    }
    
    # Eliminate x2 in R3
    aug_mat[3,] <- aug_mat[3,] - ((aug_mat[3,2]/aug_mat[2,2])*aug_mat[2,])
    
    # calculate x3 from R3
    x[3] <- aug_mat[3,4] / aug_mat[3,3]
    
    # calculate x2 from R2 with back substitution for x3
    x[2] <- (aug_mat[2,4] +  ((-1) * aug_mat[2,3]* x[3])) / aug_mat[2,2]
    
    # calculate x1 from R1 with back substitution for x2 and x3
    x[1] <- (aug_mat[1,4] +  ((-1) * aug_mat[1,2]* x[2]) + ((-1) * aug_mat[1,3]* x[3])) / aug_mat[1,1]
    
    # return the vector containing solved x1, x2 and x3
    return(x)
}

# test with given problem using both solve_eq and solve
a <- matrix(c(1,2,-1,1,-1,-2,3,5,4), nrow = 3, ncol=3)
b <- c(1,2,6)
solve_eq(a,b)
solve(a,b)

# test with equations that give zero pivot for R2  using both solve_eq and solve
a <- matrix(c(1,1,1,1,1,-1,1,-1,1), nrow = 3, ncol=3)
b <- c(7,5,3)
solve_eq(a,b)
solve(a,b)

# test with equations that give zero pivot for both R2 and R3 using both solve_eq and solve
a <- matrix(c(1,1,1,1,1,-1,-1,-1,1), nrow = 3, ncol=3)
b <- c(7,5,3)
solve_eq(a,b)
#solve(a,b)


```

