---
title: "Data605 - Assignment11"
author: "Kishore Prasad"
date: "Nov 6, 2016"
output: html_document
---

##Problem set 1

Using R's lm function, perform regression analysis and measure the significance of the independent variables for the following two data sets. In the first case, you are evaluating the statement that we hear that Maximum Heart Rate of a person is related to their age by the following equation:

MaxHR = 220 - Age

You have been given the following sample:

Age     18 23 25 35 65 54 34 56 72 19 23 42 18 39 37

MaxHR   202 186 187 180 156 169 174 172 153 199 193 174 198 183 178


Perform a linear regression analysis fitting the Max Heart Rate to Age using the lm function in R. What is the resulting equation? Is the effect of Age on Max HR significant? What is the significance level? Please also plot the fitted relationship between Max HR and Age.

**Answer:**

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(knitr)

Age <- c(18, 23, 25, 35, 65, 54, 34, 56, 72, 19, 23, 42, 18, 39, 37)
MaxHR <- c(202, 186, 187, 180, 156, 169, 174, 172, 153, 199, 193, 174, 198, 183, 178)

data <- data.frame(Age, MaxHR)

output <- summary(lm(MaxHR ~ Age, data=data))

output

ggplot(data, aes(x=Age, y=MaxHR)) + 
  geom_point() +
  stat_smooth(method=lm)

```


The resulting equation can be given by the following: \newline

$$MaxHR = `r output$coefficients[1]` - `r output$coefficients[2]` * Age$$

\newline

The effect of Age on MaxHR is quite significant. The significance value is `r output$coefficients[8]`. As we can see, this is quite close to 0 indicating that the effect is highly significant.






##Problem set 2

Using the Auto data set from Assignment 5 (also attached here) perform a Linear Regression analysis using mpg as the dependent variable and the other 4 (displacement, horse-power, weight, acceleration) as independent variables. What is the final linear regression fit equation? Which of the 4 independent variables have a significant impact on mpg? What are their corresponding significance levels? What are the standard errors on each of the coefficients? Please perform this experiment in two ways. First take any random 40 data points from the entire auto data sample and perform the linear regression fit and measure the 95% confidence intervals. Then, take the entire data set (all 392 points) and perform linear regression and measure the 95% confidence intervals. Please report the resulting fit equation, their significance values and confidence intervals for each of the two runs.


#### Part 1 - Using Full dataset


```{r, warning=FALSE, message=FALSE}

# Assuming that the data file is located in the working directory.

data <- read.table('auto-mpg.data', col.names=c('displacement', 'horsepower', 'weight', 'acceleration', 'mpg'))
head(data)

model <- lm(mpg ~ displacement + horsepower + weight + acceleration, data=data)
model

output <- summary(model)
output
```


**Answer:**

The fit equation is given by the following equation: \newline

$$mpg = `r output$coefficients[1]` + (`r output$coefficients[2]` * displacement) + (`r output$coefficients[3]` * horsepower) + (`r output$coefficients[4]` * weight) + (`r output$coefficients[5]` * acceleration) $$

\newline

The following table displays the significant variables along with the significance level, standard errors and 95% confidence interval for each of the significant variables:

```{r, warning=FALSE, message=FALSE}


vars <- c('weight', 'horsepower')
significance <- c(output$coefficients[19], output$coefficients[18])
std.err <- c(output$coefficients[9], output$coefficients[8])
confInt_95 <-  rbind(confint(model, 'weight', level=0.95)[1,], confint(model, 'horsepower', level=0.95)[1,])

sig_vars <- cbind(vars, significance, std.err, confInt_95)
kable(sig_vars)

```



#### Part 2 - Using samples


```{r, warning=FALSE, message=FALSE}

# Assuming that the data file is located in the working directory.

data <- read.table('auto-mpg.data', col.names=c('displacement', 'horsepower', 'weight', 'acceleration', 'mpg'))

set.seed(5)
sample_data <- data[sample(1:length(data$displacement), 40, replace=F),]

model <- lm(mpg ~ displacement + horsepower + weight + acceleration, data=sample_data)
model

output <- summary(model)
output
```


**Answer:**

The fit equation is given by the following equation: \newline

$$mpg = `r output$coefficients[1]` + (`r output$coefficients[2]` * displacement) + (`r output$coefficients[3]` * horsepower) + (`r output$coefficients[4]` * weight) + (`r output$coefficients[5]` * acceleration) $$

\newline

In this run (with a small sample), we find that there are no significant variables.



