---
title: "Data605 - Assignment13"
author: "Kishore Prasad"
date: "Nov 19, 2016"
output: html_document
---

##Problem set 1

Write a program to compute the derivative of $f(x) = x^3 + 2x^2$ at any value of $x$. Your function should take in a value of $x$ and return back an approximation to the derivative of $f(x)$ evaluated at that value. You should not use the analytical form of the derivative to compute it. Instead, you should compute this approximation using limits.

**Answer:**


```{r, warning=FALSE, message=FALSE}

# calculate derivative using limits

my_deriv_x3_2x2 <- function(limit, x){
    
    x1 <- x
    x2 <- x + limit
    
    delta <- (x2**3 + 2*(x2**2)) - (x1**3 + 2*(x1**2))

    return(delta/limit)
}

# check with decreasing levels of limit
my_deriv_x3_2x2(.00001, 3)
my_deriv_x3_2x2(.000001, 3)
my_deriv_x3_2x2(.0000001, 3)

# The derivative calculated numerically (3x^2 +4x) is
3*(3**2) + 4*3


# Check derivative using R built in function
dx <- deriv(~ x^3 + 2 * x^2, "x")
x <- 3
eval(dx)


```



Now, write a program to compute the area under the curve for the function $3x^2+4x$ in the range x = [1, 3]. You should first split the range into many small intervals using some really small delta x value (say 1e-6) and then compute the approximation to the area under the curve. 

**Answer:**


```{r, warning=FALSE, message=FALSE}

# calculate integration with small delta values

x <- seq(1, 3, by=1e-6)
f_x <- 3*(x**2) + 4*x
f_x_area <- f_x*1e-6
sum(f_x_area)


# The integration calculated numerically
(3**3 + 2*(3**2)) - (1**3 + 2*(1**2))

# Check integration using R built in function
integrand <- function(x) {3*(x**2) + 4*x}
integrate(integrand, lower = 1, upper = 3)

```


---


Use integration by parts to solve for 

$$\int sin(x) cos(x) dx$$

**Answer:**

Using the formula : 

$d(uv)=udv+vdu$,

$\int d(uv)= \int udv + \int vdu$,

Rearranging this gives: 

$\int udv = uv - \int vdu$


Let: $u = sin(x),  dv = cos(x) dx$

Then: 

$$\int sin(x) cos(x) dx = sin^2(x) - \int cos(x) sin(x) dx$$

Applying the power rule: $\frac{u^{1+1}}{1+1}$

substitute back $u= sin(x)$

$\frac{sin^{1+1}(x)}{1+1}$  

$$\int sin(x)cos(x)dx = \frac{sin^2(x)}{2}+C$$  


---



Use integration by parts to solve for $$\int x^2 e^x dx$$

**Answer:**

Applying the same methodology as above problem:

$$\int f(x) g'(x) dx = f(x) g(x) - \int f'(x) g(x) dx$$


$$\int x^2 e^x dx = x^2 e^x - \int 2x e^x dx$$ 

$$\int x^2 e^x dx = x^2 e^x - 2x e^x + \int 2e^x dx$$ 

$$\int x^2 e^x dx = x^2 e^x - 2x e^x + 2e^x$$


***



What is $$\frac{d}{dx}(x cos(x))$$?

**Answer:**

$$\frac{d}{dx}(x cos(x)) = cos(x) - x sin(x)$$


***


What is $$\frac{d}{dx}({e^x}^4)$$?

**Answer:**


Using chain rule: 

$\frac{df(u)}{dx} = \frac{df}{du}\cdot\frac{du}{dx}$   

$x^4 = u$

$=\frac{d}{du}(e^u)\frac{d}{dx}(x^4)$   

$\frac{d}{du}(e^u) = e^u$   

$\frac{d}{dx}(x^4) = 4x^3$   

$= e^u4x^3$    

Substituting $u = x^4$ 

$$\frac{d}{dx}(e^{x^4}) = e^{x^4}4x^3$$


