---
title: "Data605 - Assignment4"
author: "Kishore Prasad"
date: "September 17, 2016"
output: html_document
---


##Problem set 1

In this problem, we'll verify using R that SVD and Eigenvalues are related as worked out in the weekly module. Given a 3 by 2 matrix A

$A = \begin{bmatrix} 1 & 2 & 3 \\ -1 & 0 & 4 \end{bmatrix}$


write code in R to compute X = AA^T and Y = A^TA. Then, compute the eigenvalues and eigenvectors of X and Y using the built-in commands in R. Then, compute the left-singular, singular values, and right-singular vectors of A using the svd command. Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y. In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the non-zero singular values of A.

Your code should compute all these vectors and scalars and store them in variables. Please add enough comments in your code to show me how to interpret your steps.


**Answer:**


```{r, warning=FALSE, message=FALSE}

# Define the matrix
A <- matrix(c(1,2,3,-1,0,4), nrow = 2, ncol = 3, byrow = TRUE)
A

# Calculate X as AAT
X <- A %*% t(A)
X

#Calculate Y as ATA
Y <- t(A) %*% A
Y

#Eigenvalues and Eigenvectors of X rounded to 2 decimals

Eigen_Values_X <- round(unlist(eigen(X)[1]),2)
Eigen_Values_X

Eigen_Vector_X <- matrix(round(unlist(eigen(X)[2]),2), nrow=2, ncol=2, byrow=FALSE)
Eigen_Vector_X


#Eigenvalues and Eigenvectors of Y rounded to 2 decimals

Eigen_Values_Y <- round(unlist(eigen(Y)[1]),2)
Eigen_Values_Y

Eigen_Vector_Y <- matrix(round(unlist(eigen(Y)[2]),2), nrow=3, ncol=3, byrow=FALSE)
Eigen_Vector_Y

# Singular values, Left-singular and Right-singular vectors of A

Singular_A <- round(svd(A)$d,2)
Singular_A

Left_sing_A <- round(svd(A)$u,2)
Left_sing_A

Right_sing_A <- round(svd(A)$v,2)
Right_sing_A


# Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y. 

# comparing the Left singular with eigen vector of X
Left_sing_A
Eigen_Vector_X

# comparing the Right singular with eigen vector of Y
Right_sing_A
Eigen_Vector_Y


# In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the non-zero singular values of A.

# Singular values of A
Singular_A

# Square of the Singular Values to compare with eigen values of X and Y
Singular_A^2

# Eigen Values of X is the same as the square of the singular values of A
Eigen_Values_X

# Eigen Values of Y  is the same as the square of the singular values of A
Eigen_Values_Y

```




##Problem set 1


Using the procedure outlined in section 1 of the weekly handout, write a function to compute the inverse of a well-conditioned full-rank square matrix using co-factors. In order to compute the co-factors, you may use built-in commands to compute the determinant. Your function should have the following signature:

B = myinverse(A)

where A is a matrix and B is its inverse and A x B = I. The off-diagonal elements of I should be close to zero, if not zero. Likewise, the diagonal elements should be close to 1, if not 1. Small numerical precision errors are acceptable but the function myinverse should be correct and must use co-factors and determinant of A to compute the inverse. 

**Answer:**

```{r, warning=FALSE, message=FALSE}

cofactors <- function(A) {
    dim_a <- nrow(A)                # Determine the dimension of A
    orig_A <- A                     # Store the original A matrix

    cofactor_matrix <- diag(dim_a)
    
    # set up a loop for rows from 1 to dim_a
    for(i in 1:dim_a) {
        # set up loops for the respective columns 
        for(j in 1:dim_a){
            # Calculate the cofactor for the element i, j 
            cofactor_matrix[i,j] <- det(A[-i,-j]) * ((-1)^(i+j))
#            cofactor_matrix[i,j] <- det(A[-i,-j]) 
        }
    }
    return(cofactor_matrix)
}


inverse <- function(A) {
    # calculate the determinant of A
    det_A <- det(A)
    
    # calculate the cofactor of A
    cofactor_A <- cofactors(A)

    # calculate the inverse
    inverse_A <- t(cofactor_A) / det_A  

    return(inverse_A)
}

# Define a 3x3 Matrix
A <- matrix(c(1,2,3,-1,0,4, 2,5,7), nrow = 3, ncol = 3, byrow = TRUE)
A

# calculate the inverse rounded to 2 decimals.
B <- round(inverse(A),2)
B

# The multiplication of A * B with zero decimal places should give the identity matrix.
round(A %*% B, 0)

# Try with 4x4 matrix
A <- matrix(c(1, 4, 2, -1, 0, 4, 2, 5, 7, 12, 8, 4, 3, 6, 4, 2), nrow = 4, ncol = 4, byrow = TRUE)
A
B <- round(inverse(A),2)
B
round(A %*% B, 0)

```

